Higgs Assisted Razor Search for Higgsinos at 100 TeV

### 2019-07-20 main.tex
Addressing more comments

\abstract{% 
  A 100 TeV proton-proton collider will be an extremely effective 
  way to probe the electroweak sector of the Minimal Supersymmetric 
  Standard Model (MSSM). In this paper, we describe a search strategy  
  for discovering pair-produced Higgsino-like next-to-lightest supersymmetric particles (NLSP)  at a 100 TeV 
  hadron collider that decay to Bino-like lightest supersymmetric particle (LSP) via intermediate $Z$ and 
  SM Higgs bosons that in turn decay to a pair of leptons and a pair of 
  $b$-quarks respectively: $\N_{2}\N_{3}
  \rightarrow (Z\N_1)(h\N_1)\rightarrow 
  bb\ell\ell+\N_1\N_1$. In addition, we examine 
  the potential for machine learning techniques to boost the power of 
  our searches. Using this analysis, Higgsinos up to 1.3 TeV can be 
  discovered for Binos  up to 0.9 TeV using 3000 fb$^{-1}$ of data,  
  Additionally, Higgsinos up  to 1.8 TeV can be excluded for Binos up 
  to 1.3 TeV at 95\% C.L.   This search channel extends the multi-lepton search limits, especially in the region where the mass difference between
Higgsino NLSP and Bino LSP is smaller. \Shufang{Update the numbers.}
}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.49\textwidth]{images/xsection_plot.pdf}
  \includegraphics[width=0.49\textwidth]{images/higgsino_br_plot.png}
  \caption{Higgsino pair production $\widetilde\chi_2^0\widetilde\chi_3^0$ cross section (left panel), calculated using Prospino2~\cite{Beenakker:1999xh},  and decay decay branching fractions (right panel), calculated in SUSY-HIT~\cite{Djouadi:2006bz},  as a function of $\mu$ for $M_1=25$ GeV. However, it
  should be noted that the dependence on $M_1$ is very weak.   \Shufang{How about Br? What's the vcalue of M1 used?}\\
  \Shufang{Could you make the two plot same aspect ratios? Using plot digitizer.  Also, for ylabel in left panel, the tilde should not go above 0.  Just above $\chi$.  Is there way to change the xlabel of right panel to be $\mu$ instead of $|\mu|$?  Need to be consistent with the xlabels.}}
  \label{fig:xsection_plot}
\end{figure}

@@

Since we expect our signal process to have a dilepton resonance from an on-shell
$Z$ boson, we restricted the phase space for event generation for backgrounds to
the region where the invariant mass of dilepton pairs lies between 80 and 100
GeV. Additionally, the Bino dark matter that escapes the detector would result
in a large amount of missing transverse energy ($\met$), so we required
a minimum $\met$ of 100 GeV at the parton level for the backgrounds as
well.  \Shufang{This is only for BG, right?  No such cuts are imposed for signal generation?}

@@

  \item \emph{Razor variables}:  Two razor variables are used in our analyses: 
\begin{align}
M_R &= \sqrt{(E_Z+E_h)^2 - (p_Z^z + p_h^z)^2},\\
M_T^R &= \sqrt{\frac{1}{2}\left[\slashed{E}_T(|\vec{p}_{ZT}|+|\vec{p}_{hT}|)
- \vec{\slashed{E}}_T\cdot(\vec{p}_{ZT}+\vec{p}_{hT})\right]}
\end{align}
 for $E_{Z,h}$ and $\vec{p}_{Z,h}$ being the reconstructed energy and momentum for $Z$ and $h$, respectively.
      Representative kinematic
  distributions of these variables for signal events of a 1 TeV Higgsino and a 25 GeV Bino, as well as $tt$ and $tbW$ backgrounds are
  shown in Fig.~\ref{fig:razor_histos}, after the trigger, identification, and
  $\slashed{E}_{T}$ cuts.  \Shufang{How about mll cut? BG is generated with mll cut.}  The
  distribution for the \emph{bbWW} background is not shown, since it is negligible in
  comparison to the others.   We can see that the distribution of $M_R$ is peaked
  around 1 TeV for the signal, which is consistent with what we would expect as   it corresponds to the mass difference between the NLSP and LSP.  The $tt$ background peaks around 500 GeV, while the distribution of $tbW$ background almost overlaps with that of the signal process.  Similar behaviour also appears for $M_T^R$ distributions.  
  For this mass combination, requiring $M_R >$ 800
  GeV and $M_T^R >$ 400 GeV yields the greatest significance.

@@

\begin{table}[h]
  \centering
  \input{tables/cutflowtable.tex}
  \caption{Representative cut flow table for the benchmark point $\mu=1$ TeV,
    $M_1 = 25$ GeV at 100  TeV $pp$ collider, for a traditional cut-and-count analysis. All cross sections
    are given in unit of fb, and the units for the missing energy, invariant
    mass, and razor variable cuts are GeV. The significance, $S/\sqrt{B}$, is
    calculated for an integrated luminosity of 3 ab$^{-1}$.   \Shufang{MET $>$ 400 or $<$ 400? }}
\label{tab:cc_cutflowtable}
\end{table}

@@

For each signal mass combination and each background process, we preselected
events that passed the lepton trigger, contained two SFOS leptons and two
$b$-tagged jets, and calculated a number of kinematic variables for each event. \Shufang{Again, generator level MET, mll cut for the BG?} We then
placed these kinematic variables in an array, with each row corresponding to an event, and
the columns corresponding to the kinematic variables. A mixture of low-level and high level
kinematic variables was shown to have the greatest effectiveness. The kinematic variables chosen were
$m_{\ell\ell}$, $m_{bb}$, $M_R$, $M_T^R$, $\slashed{E}_T,$ the total hadronic
transverse energy $H_T$,   and transverse momenta of individual final state
leptons and $b$-quarks: $p_T^{\ell_1}$, $p_T^{\ell_2}$, $p_T^{b_1}$, and $p_T^{b_2}$. We
then divided the generated events into training and test sets, with the training set
comprising 75\% of signal events and 30\% of background events.  We used the
\texttt{scikit-learn} package  \cite{Pedregosa2011} to implement our analysis.
A boosted decision tree classifier was trained with 1000 weak learners and a
learning rate of 0.025.  
After the classifier is trained, we use it to assign
scores to individual events. A more negative score indicates a more
background-like event, and a more positive score denotes a more signal-like
event. After the scores have been assigned to the events in the test sets, we
can use this score just as we would a regular kinematical variable, that is, we
apply a cut that selects events with a minimum of this score. The value of the
cut is chosen to maximize the significance for each signal benchmark point. The
distribution of scores for the specific benchmark point ($\mu$ = 1 TeV, $M_1$
= 25 GeV) is shown in figure \ref{fig:bdt_response}. We observe that there is an
appreciable separation between the signal and background distributions.

@@

\caption{ Discovery (red) and exclusion (blue) contours for the traditional
  cut-and-count analysis (solid) and boosted decision tree analysis (dashed),
  for 100 TeV $pp$ collider with an integrated luminosity of 3000 fb$^{-1}$.
\Shufang{Can we change the xlabel of all plots to $\mu$ instead of $|\mu|$ or it is difficult to do?}\\  \Shufang{Did you save the data for this plot?  It is actually not difficult to overlay ref 10 results.  They only have a few data points and can easily be read out using plot-digitizer.  By comparison, we can clearly see that our reach is better for large M1 region.}
}

@@

Fig.~\ref{fig:contours} shows the expected reach  for 95\%
C.L. exclusion and 5$\sigma$ discovery in the parameter space of $\mu$ vs.
$M_1$ for 100 TeV $pp$ collider with an integrated luminosity of 3000
fb$^{-1}$. The cut-and-count strategy is able to discover Higgsinos up to 750
GeV  and exclude them up to 1.3 TeV. \Shufang{Cross check the numbbers. Not entirely consistent with the plot.}    The boosted decision tree analysis, improves this reach -- it
is able to discover Higgsinos up to 1.3 TeV, and exclude them up to 1.8 TeV in
favorable points in parameter space.  Similarly, Binos can be discovered up to
300 GeV and excluded up to 800 GeV with the rectangular cut analysis, and the
BDT analysis can discover them up to about 900 GeV and exclude them up to 1.3
TeV. Since the background left after cut is relatively low, We do not include
systematic errors in these estimates.

@@

Overall, we find that the reach of our analysis strategy is a significant
improvement over that of the LHC.    We found that the rectangular cut strategy has
the potential to discover Higgsino LSPs up to a mass of 750 GeV, and exclude
them up to a mass of 1.3 TeV. As expected, the boosted decision tree classifier
performs better, with the ability to discover Higgsino NLSPs up to 1.3 TeV and
exclude them up to a mass of 1.8 TeV. With the BDT analysis, Bino dark matter
candidates can be discovered up to 900 GeV, and excluded up to 1.3 TeV. In the
process, we highlighted the importance of generating enough Monte Carlo events
to estimate the huge backgrounds at a 100 TeV $pp$ collider.  \Shufang{Update the numbers.}

>>>>

\abstract{%

  A 100 TeV proton-proton collider will be an extremely effective way to probe
  the electroweak sector of the Minimal Supersymmetric Standard Model (MSSM).
  In this paper, we describe a search strategy  for discovering pair-produced
  Higgsino-like next-to-lightest supersymmetric particles (NLSP)  at a 100 TeV
  hadron collider that decay to Bino-like lightest supersymmetric particle
  (LSP) via intermediate $Z$ and SM Higgs bosons that in turn decay to a pair
  of leptons and a pair of $b$-quarks respectively: $\N_{2}\N_{3} \rightarrow
  (Z\N_1)(h\N_1)\rightarrow bb\ell\ell+\N_1\N_1$. In addition, we examine the
  potential for machine learning techniques to boost the power of our searches.
  Using this analysis, Higgsinos up to 1.4 TeV can be discovered for Binos with
  masses of about
  0.9 TeV using 3000 fb$^{-1}$ of data. Additionally, Higgsinos up to 1.8
  TeV can be excluded for Binos with masses of about 1.3 TeV at 95\% C.L.   This search
  channel extends the multi-lepton search limits, especially in the region
  where the mass difference between Higgsino NLSP and Bino LSP is smaller.
\Shufang{Update the numbers.} \Adarsh{They are consistent with the text now.}}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.49\textwidth]{images/xsection_plot.pdf}
  \includegraphics[width=0.49\textwidth]{images/br_plot.pdf}
  \caption{Higgsino pair production $\widetilde\chi_2^0\widetilde\chi_3^0$
    cross section (left panel), calculated using
    Prospino2~\cite{Beenakker:1999xh},  and decay decay branching fractions
    (right panel), calculated in SUSY-HIT~\cite{Djouadi:2006bz},  as a function
    of $|\mu|$ for $M_1=25$ GeV. However, it should be noted that the dependence
    on $M_1$ is very weak. \Shufang{How about Br? What's the vcalue of M1
    used? Could you make the two plot same aspect ratios? Using plot
digitizer. Also, for ylabel in left panel, the tilde should not go above 0.
Just above $\chi$.  Is there way to change the xlabel of right panel to be
$\mu$ instead of $|\mu|$?  Need to be consistent with the xlabels.
}\Adarsh{The value of $M_1$ used is 25 GeV for the branching ratio plot as
well. I was able to rerun prospino and susyhit, so I didn't need to run the
plotdigitizer for these plots. I also changed the tildes as you suggested.
I changed the x-axis of both plots to $|\mu|$ for consistency - is that ok?}
}
  \label{fig:xsection_plot}
\end{figure}

@@

Since we expect our signal process to have a dilepton resonance from an on-shell
$Z$ boson, we restricted the phase space for event generation for backgrounds to
the region where the invariant mass of dilepton pairs lies between 80 and 100
GeV. Additionally, the Bino dark matter that escapes the detector would result
in a large amount of missing transverse energy ($\met$), so we required
a minimum $\met$ of 100 GeV at the parton level for the backgrounds as
well. \Shufang{This is only for BG, right?  No such cuts are imposed for
signal generation?} \Adarsh{Yes, the MET cut was only for background.}

@@

  \item \emph{Razor variables}:  Two razor variables are used in our analyses: 
\begin{align}
M_R &= \sqrt{(E_Z+E_h)^2 - (p_Z^z + p_h^z)^2},\\
M_T^R &= \sqrt{\frac{1}{2}\left[\slashed{E}_T(|\vec{p}_{ZT}|+|\vec{p}_{hT}|)
- \vec{\slashed{E}}_T\cdot(\vec{p}_{ZT}+\vec{p}_{hT})\right]}
\end{align}
 for $E_{Z,h}$ and $\vec{p}_{Z,h}$ being the reconstructed energy and momentum for $Z$ and $h$, respectively.
      Representative kinematic
  distributions of these variables for signal events of a 1 TeV Higgsino and a 25 GeV Bino, as well as $tt$ and $tbW$ backgrounds are
  shown in Fig.~\ref{fig:razor_histos}, after the trigger, identification,
  $m_{\ell^+\ell^-}$, and
  $\slashed{E}_{T}$ cuts.  \Shufang{How about mll cut? BG is generated with mll
  cut.}  \Adarsh{I have added the $m_{ll}$ cut to the sentence.} The
  distribution for the \emph{bbWW} background is not shown, since it is negligible in
  comparison to the others.   We can see that the distribution of $M_R$ is peaked
  around 1 TeV for the signal, which is consistent with what we would expect as   it corresponds to the mass difference between the NLSP and LSP.  The $tt$ background peaks around 500 GeV, while the distribution of $tbW$ background almost overlaps with that of the signal process.  Similar behaviour also appears for $M_T^R$ distributions.  
  For this mass combination, requiring $M_R >$ 800
  GeV and $M_T^R >$ 400 GeV yields the greatest significance.

@@

\begin{table}[h]
  \centering
  \input{tables/cutflowtable.tex}
  \caption{Representative cut flow table for the benchmark point $\mu=1$ TeV,
    $M_1 = 25$ GeV at 100  TeV $pp$ collider, for a traditional cut-and-count analysis. All cross sections
    are given in unit of fb, and the units for the missing energy, invariant
    mass, and razor variable cuts are GeV. The significance, $S/\sqrt{B}$, is
    calculated for an integrated luminosity of 3 ab$^{-1}$. \Shufang{MET $>$
  400 or $<$ 400? } \Adarsh{$> 400$ - I've fixed it now.}}
\label{tab:cc_cutflowtable}
\end{table}

@@

For each signal mass combination and each background process, we preselected
events that passed the lepton trigger, contained two SFOS leptons and two
$b$-tagged jets, and calculated a number of kinematic variables for each event.
\Shufang{Again, generator level MET, mll cut for the BG?} \Adarsh{Yes.} We then
placed these kinematic variables in an array, with each row corresponding to an
event, and the columns corresponding to the kinematic variables. A mixture of
low-level and high level kinematic variables was shown to have the greatest
effectiveness. The kinematic variables chosen were $m_{\ell\ell}$, $m_{bb}$,
$M_R$, $M_T^R$, $\slashed{E}_T,$ the total hadronic transverse energy $H_T$,
and transverse momenta of individual final state leptons and $b$-quarks:
$p_T^{\ell_1}$, $p_T^{\ell_2}$, $p_T^{b_1}$, and $p_T^{b_2}$. We then divided
the generated events into training and test sets, with the training set
comprising 75\% of signal events and 30\% of background events.  We used the
\texttt{scikit-learn} package  \cite{Pedregosa2011} to implement our analysis.
A boosted decision tree classifier was trained with 1000 weak learners and a
learning rate of 0.025.

After the classifier is trained, we use it to assign scores to individual
events. A more negative score indicates a more background-like event, and a
more positive score denotes a more signal-like event. After the scores have
been assigned to the events in the test sets, we can use this score just as we
would a regular kinematical variable, that is, we apply a cut that selects
events with a minimum of this score. The value of the cut is chosen to maximize
the significance for each signal benchmark point. The distribution of scores
for the specific benchmark point ($\mu$ = 1 TeV, $M_1$ = 25 GeV) is shown in
figure \ref{fig:bdt_response}. We observe that there is an appreciable
separation between the signal and background distributions.

@@

\caption{ Discovery (red) and exclusion (blue) contours for the traditional
  cut-and-count analysis (solid) and boosted decision tree analysis (dashed),
  for 100 TeV $pp$ collider with an integrated luminosity of 3000 fb$^{-1}$.
\Shufang{Can we change the xlabel of all plots to $\mu$ instead of $|\mu|$ or
it is difficult to do? Did you save the data for this plot?  It is
actually not difficult to overlay ref 10 results.  They only have a few data
points and can easily be read out using plot-digitizer.  By comparison, we can
clearly see that our reach is better for large M1 region.} \Adarsh{I did not
save the data for this plot - I thought I could always regenerate it - however,
the HPC folks changed their policy on the number of files per GB that they
would allow - MadGraph and MadEvent produce a lot of files - for the Exotic
Higgs project, I came up with ways to deal with this, but to do the same for
this project would take a very large amount of time. I could change the xlabel
to $\mu$ instead of $|\mu|$, but isn't $|\mu|$ more accurate, since $\mu$ can
be positive or negative?}
}

@@

Fig.~\ref{fig:contours} shows the expected reach  for 95\% C.L. exclusion and
5$\sigma$ discovery in the parameter space of $\mu$ vs.  $M_1$ for 100 TeV $pp$
collider with an integrated luminosity of 3000 fb$^{-1}$. The cut-and-count
strategy is able to discover Higgsinos up to 850 GeV  and exclude them up to
1.3 TeV. \Shufang{Cross check the numbers. Not entirely consistent with the
plot.} \Adarsh{I updated the numbers - how do they look?} The boosted decision
tree analysis, improves this reach -- it is able to discover Higgsinos up to
1.4 TeV, and exclude them up to 1.8 TeV in favorable points in parameter space.
Similarly, Binos can be discovered up to 300 GeV and excluded up to 700 GeV
with the rectangular cut analysis, and the BDT analysis can discover them up to
about 900 GeV and exclude them up to 1.3 TeV. Since the background left after
cut is relatively low, we do not include systematic errors in these estimates.

@@

Overall, we find that the reach of our analysis strategy is a significant
improvement over that of the LHC. We found that the rectangular cut strategy has
the potential to discover Higgsino LSPs up to a mass of 850 GeV, and exclude
them up to a mass of 1.3 TeV, for bino masses around 350 GeV and 700 GeV
respectively. 
As expected, the boosted decision tree classifier
performs better, with the ability to discover Higgsino NLSPs up to 1.4 TeV and
exclude them up to a mass of 1.8 TeV, for bino masses of about 900 GeV and 1.3
TeV respectively. With the BDT analysis, Bino dark matter
candidates can be discovered up to 0.9 TeV, and excluded up to 1.3 TeV. In the
process, we highlighted the importance of generating enough Monte Carlo events
to estimate the huge backgrounds at a 100 TeV $pp$ collider.  \Shufang{Update
the numbers.} \Adarsh{They are updated now.}
